Against the Planetary Interpretation of PTFO$\,$8-8695b

### 2020-04-02 paper/ms.tex
paper edits

@@

% TODO:
% Figure: CDIPS LC + SPOC LC (left two panels). Right panel:
% periodogram.
%

@@

@@

\paragraph{Fourier analysis}
Figure~\ref{fig:splitsignal} could also be given in the frequency
domain, rather than the time domain.
The largest peaks in 
the Lomb-Scargle periodogram (CITE) of the raw light-curve are
at $P_{\rm s}$ and $P_{\rm \ell}$.
Both peaks of course have aliases.
The $P_{\rm \ell}$ peak has larger ampitude.
Subtracting out the long-period signal, the short-period signal
dominates the periodogram, and vice-versa.
The periodogram of the final residual (Figure~\ref{fig:splitsignal}
bottom row) shows a weakly significant, poorly resolved peak at
$\approx$8 days, consistent with the visual impression in the time
domain that there could be a weak long-period signal present.

>>>>

% \subsection{Observations}

PTFO 8-8695 was observed by TESS with Camera 1, CCD 1, from December
15, 2018 to January 6, 2019, during the sixth sector of science
operations.  The star is designated TIC 264461976 in the TESS Input
Catalog \citep{stassun_TIC_2018}.  The pixel data for an $11\times11$
array surrounding PTFO 8-8695 were averaged into 2-minute stacks by
the onboard computer. 
Each 2048$\times$2048 image was also averaged into 30-minute stacks,
and saved as a ``full frame image'' (FFI).

%FIXME these paragraphs are directly copied from WASP-4 paperI
%FIXME these paragraphs are directly copied from WASP-4 paperI
The data were downlinked via the Deep Space
Network\footnote{\url{deepspace.jpl.nasa.gov}}, and the
spacecraft timestamps were calibrated against the ground-station
clocks.  The spacecraft clock times were then transformed by the
Payload Operations Center into the {\it Temps Dynamique Barycentrique}
(TDB) reference system.  The images were then reduced to lightcurves
by the Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.

%FIXME these paragraphs are directly copied from WASP-4 paperI
%FIXME these paragraphs are directly copied from WASP-4 paperI
We began our analysis with the Presearch Data Conditioning (PDC)
lightcurve, which has had non-astrophysical variability removed
through the methods discussed by \citet{smith_kepler_apertures_2017}
and \citet{smith_kepler_PDC_2017}.  We then processed the lightcurve
as follows.  First, we removed all points with non-zero quality flags.
This removed data that might have been adversely affected by
``momentum dumps,'' the firing of thrusters and resetting of reaction
wheels\footnote{The spacecraft pointing and momentum dumps are
described in the data release notes:
\url{archive.stsci.edu/missions/tess/doc/tess_drn/tess_sector_07_drn09_v03.pdf}}
that took place every 2.5 days during sector 2.  The data during these
events were assigned quality flags corresponding to ``Reaction Wheel
Desaturation Event'' and ``Manual Exclude'' \citep[][Table
28]{tess_data_product_description_2018}.  For WASP-4, these flags were
simultaneously set for 54 distinct cadences, and there were 10
momentum dumps, averaging about 10 minutes of flagged data per dump.
Out of caution, we clipped out an additional 10 minutes before and
after every momentum dump.  We also removed the data within the first
and last hours of both orbits, because of correlated red noise that
appears during those time ranges.

%FIXME these paragraphs are directly copied from WASP-4 paperI
%FIXME these paragraphs are directly copied from WASP-4 paperI
All told, we removed 8\% of the original data points, and were left
with 18{,}165 measurements of the relative flux of WASP-4.  We
normalized the data by dividing out the median flux.  We converted the
timestamps from BTJD$_{\rm TDB}$ into BJD$_{\rm TDB}$ by adding the
appropriate 2{,}457{,}000 day offset
\citep{tess_data_product_description_2018}.  Many of these and
subsequent processing steps were performed using
\texttt{astrobase}~\citep{bhatti_astrobase_2018}. We did not
``flatten'' the lightcurves, as is often done with splines,
polynomials, or Gaussian processes.  Instead, we modeled the
out-of-transit flux variations simultaneously with the transit
parameters, as described below.

% \subsection{Star and planet parameters}
% \label{sec:system_parameters}
% 
% \begin{deluxetable}{lccc}
% \tabletypesize{\scriptsize}
% \tablecaption{Selected system parameters of WASP-4b\label{tbl:params}}
% \tablenum{1}
% 
% \tablehead{
% \colhead{Parameter} & \colhead{Value} & \colhead{68\% Confidence Interval} & \colhead{Comment}
% }
% 
% \startdata
% {\it Transit/RV parameters:} & & & \\
%   $R_{\rm p}/R_\star$                        & $0.15201$              & $+0.00040$, $-0.00033$      & A \\
%   $i$~[deg]                                  & $89.06$                & $+0.65$, $-0.84$            & A \\
%   $a/R_\star$                                & $5.451$                & $+0.023$, $-0.052$          & A \\
%   $u_{\rm linear}$                           & $0.382$                & ---                         & A \\
%   $u_{\rm quad}$                             & $0.210$                & ---                         & A \\
%   $K$~[m~s$^{-1}$]                           & $241.1$                & $+2.8$, $-3.1$              & B \\
% {\it Stellar parameters:} & & & \\
%   $T_{\rm eff}$~[K]                          & $5400$                 & $\pm 90$                    & C \\
%   $\log g_\star$~[cgs]                       & $4.47$                 & $\pm 0.11$                  & C \\
%   $[{\rm Fe/H}]$                             & $-0.07$                & $\pm 0.19$                  & C \\
%   $F_{\rm bol}$~[erg~cm$^{-2}$~s$^{-1}$]     & $2.802\times10^{-10}$  & $\pm 0.076\times10^{-10}$   & D \\
%   $A_V$~[mag]                                & $0.03$                 & $+0.02, -0.01$              & D \\
%   $\pi$~[mas]                                & $3.7145$               & $0.0517$                    & F \\
%   $R_\star$~[R$_{\odot}$]                    & $0.893$                & $\pm 0.034$                 & E \\
%   $\rho_\star$~[g~cm$^{-3}$]                 & $1.711$                & $+0.022$, $-0.048$          & E \\
%   $M_\star$~[M$_{\odot}$]                    & $0.864$                & $+0.084$, $-0.090$          & E \\
%   $T$ magnitude                              & $11.778$               & $\pm 0.018$                 & G \\
% {\it Planetary parameters:} & & & \\
%   $a$~[AU]                                   & $0.0226$               & $+0.0007$, $-0.0008$        & E \\
%   $M_p$~[M$_{\rm Jup}$]                      & $1.186$                & $+0.090$, $-0.098$          & E \\
%   $R_p$~[R$_{\rm Jup}$]                      & $1.321$                & $\pm 0.039$                 & E \\
% \enddata
% 
% \tablecomments{
%   (A) From phase-folded TESS lightcurve (\S~\ref{sec:measurement}).
%   Orbital periods are in Table~4. The limb darkening parameters were
%   allowed to float around the \citet{claret_limb_2017} prediction, but
%   were unconstrained.
%   (B) \citet{triaud_spin-orbit_2010}.
%   (C) From HARPS spectra \citep{doyle_accurate_2013}.
%   (D) \citet{stassun_accurate_2017}.
%   (E) This work, see \S~\ref{sec:system_parameters}.
%   (F) \citet{gaia_collaboration_gaia_2018}.
%   (G) \citet{stassun_TIC_2018}.
% }
% 
% \vspace{-1cm}
% \end{deluxetable}
% 
% We calculated the stellar and planetary parameters in the following
% way.  We computed the star's spectral energy distribution based on the
% {\it Gaia} DR2 parallax (after making the small correction advocated
% by \citealt{stassun_evidence_2018}) and the broadband magnitudes from
% the available all-sky catalogs: $G$ from {\it Gaia\/} DR2, $B_T$ and
% $V_T$ from {\it Tycho-2}, $BVgri$ from {\it APASS}, $JHK_S$ from {\it
% 2MASS}, and the {\it WISE}~1--4 passbands, thus spanning the
% wavelength range 0.4--22~$\mu$m.
% We adopted the effective temperature
% from the work by \citet{doyle_accurate_2013}, who determined the
% spectroscopic parameters of WASP-4 using high-signal-to-noise
% observations with the High Accuracy Radial-velocity Planet Searcher
% (HARPS).  Then, we determined the stellar radius through the
% combination of the bolometric luminosity and the effective
% temperature, using the Stefan-Boltzmann law.  To determine the stellar
% mass, we first computed the mean stellar density based on the value of
% $a/R_\star$ that gave the best fit to the phase-folded TESS lightcurve
% (for the relevant equation, see \citealt{seager_unique_2003} or
% \citealt{winn_exoplanet_2010}).  The mass was calculated from the
% radius and density, and the orbital distance was also calculated from
% the radius and $a/R_\star$.  The planetary radius was calculated as
% the product of $R_\star$ and $R_{\rm p}/R_\star$.  Finally, the planet
% mass was calculated based on the stellar mass, the radial-velocity
% amplitude observed by \citet{triaud_spin-orbit_2010}, and the orbital
% inclination.
% 
% Table~1 gives the resulting parameters, which we adopted for the
% remaining analysis.  \added{The uncertainties in our derived stellar
% and planetary parameters are propagated according to standard analytic
% formulae, under the assumption that the variables are uncorrelated and
% normally distributed.} Our \added{system} parameters are in agreement
% with those of previous investigators, but have the benefit of
% incorporating the {\it Gaia} parallax
% \citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013,huitson_gemini_2017}.
% By comparing the star's luminosity and spectroscopic parameters with
% the outputs of the Yonsei-Yale stellar-evolutionary models, we found
% that WASP-4 is a main-sequence star with an age of approximately
% $7\,{\rm Gyr}$ \added{\citep{demarque_y2_2004}}.

@@

Our initial inspection of the light-curve, in both its 2-minute PDCSAP
and 30-minute FFI forms, revealed an obvious beat signal
(Figure~\ref{fig:splitsignal}, top panel).

As a precursor to more detailed analysis, we calculated generalized
Lomb-Scargle periodograms using \texttt{astrobase}
\citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}. 
The two largest peaks in the Lomb-Scargle periodogram of the
light-curve were clearly separated at $P_{\rm s} \approx 0.448\,{\rm
days}$ and $P_{\rm \ell} \approx 0.499\,{\rm days}$.
The $P_{\rm \ell}$ peak had the larger ampitude of the two.
Smaller harmonics from each of these dominant two signals were also
present.

Initial fitting experiments with splines showed that subtracting out
the long-period signal, the short-period signal dominated the
periodogram, and vice-versa.
However it quickly became clear that a mroe careful modelling approach
than ``flattening'' the light-curve with splines or Gaussian proceses
would be preferable, in order to preserve the power at each frequency.

@@

We opted to model the light-curve as a linear combination of Fourier
harmonics at the short and long periods, plus a transit at the
short period.
Symbolically,
\begin{equation}
  f = f_{\rm s} + f_{\rm \ell} = f_{\rm transit,s} + f_{\rm Fourier,s} + f_{\rm Fourier,\ell},
\end{equation}
where $f_{\rm s}$ is the relative flux at the short period, and
$f_{\rm \ell}$ is the flux at the long period.
Decomposing each term,
\begin{align}
  f = &f_{\rm transit,s} + \sum_{n=1}^{n=N} A_n \sin(n\omega_{\rm s}t)
  + \sum_{n=1}^{n=N} B_n \cos(n\omega_{\rm s}t)\\
  &+ \sum_{m=1}^{m=M} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
  + \sum_{m=1}^{m=M} B_m \cos([m\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
\end{align}
for $N$ and $M$ the total number of harmonics at the short and long period, respective,
$A_i$ and $B_i$ for each harmonic term, and $\omega_i = 2\pi / P_i$ the angular
frequency for $i$ the short or long period index. 
We fixed the ``phase-offset'' for the short period signal to be zero,
and let the reference time of the long period signal float by introducing
$\phi_{\rm \ell}$. 
Since we did not a priori know how many harmonics would be
appropriate, we let the number of harmonics vary, and
used the Bayesian information criterion to choose the
appropriate model.

For example, one possible model could be a transit, plus $N=2$ harmonics
of sines and cosines at the short period, plus $M=1$  harmonics at the long
period.
In this case, the free parameters would be as follows.
For the transit, we would fit for
the impact parameter, the planet-to-star radius ratio, two quadratic limb darkening parameters,
the planet orbital period (equal to the short period), the reference time
for the transit, and the mean flux.
There would be 4 additional Fourier amplitudes at the short period,
plus 2 Fourier amplitudes at the long period, and well as the long period itself
and its phase.
For this case, we therefore fitted 14 free parameters.

% 
% I gave it a pass on the PTFO data, binning from 2->10 minute sampling
% to speed up the execution time. Results from two models are below.
% Case 1 is N=2,M=1. Case 2 is N=2,M=2. Both fits are fine, and I get
% decent convergence with a few minutes each. The rotation parameters
% (phase + amplitudes) are quite correlated, so I set a reasonably
% strong prior on the rotation phase.
% 
% The main result that both models agree on: the extra power at the
% orbital frequency seems to not be ellipsoidal. Both of these models
% prefer to put the power at 1x the orbital frequency, not 2x the
% orbital frequency! This gives both Aorb0 and Borb0 non-zero, while
% Aorb1 and Borb1 seem to be more consistent with zero.

% Using the cleaned PDC lightcurve, we applied the Box Least Squares
% algorithm \citep{kovacs_box-fitting_2002} to estimate the orbital
% period, transit duration, and a reference epoch using the TESS data
% alone.  Based on the results, we isolated the data within 4 transit
% durations of each transit midpoint.  To find the transit parameters
% that best fit the data, we first fitted a line to the out-of-transit
% flux measurements surrounding each transit, and divided it out.  We
% then created a phase-folded lightcurve \added{from all 18 transits}
% and fitted a standard transit model using the analytic formulae given
% by \citet{mandel_analytic_2002} and implemented by
% \replaced{\citet{kreidberg_batman_2015}}{\citet[][\texttt{BATMAN}]{kreidberg_batman_2015}}
% We assumed the orbit to be circular, consistent with the limits from
% radial velocities and occultation timing
% \citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
% The free parameters were the reference epoch, the planet to star
% radius ratio $R_{\rm p}/R_\star$, the orbital distance to stellar
% radius ratio $a/R_\star$, the inclination $i$, two quadratic
% limb-darkening coefficients $(u_{\rm linear}, u_{\rm quad})$, and the
% orbital period $P$.
% 
% We sampled the posterior probability distribution for all the
% parameters using the algorithm proposed by
% \citet{goodman_ensemble_2010} and implemented by
% \replaced{\citet{foreman-mackey_emcee_2013}}{\citet[][\texttt{emcee}]{foreman-mackey_emcee_2013}}.
% Table~1 gives the results, which are in reasonable agreement with the
% parameters reported by {\it e.g.},
% \citet{southworth_high-precision_2009} and
% \citet{huitson_gemini_2017}.  Figure~\ref{fig:phasefold} shows the
% phase-folded lightcurve.
% 
% To measure the transit times, we returned to the `cleaned' PDC time
% series and fitted the data within four transit durations of each
% transit separately. We used four free parameters: the time of
% mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
% slope and intercept of a linear trend to account for any slow
% variations unrelated to the transit.  We fixed the remaining
% parameters at the values that had been determined from the
% phase-folded TESS lightcurve.  The uncertainty in each
% \added{photometric} data point was set equal to the root-mean-square
% (rms) level of the out-of-transit data.
% 
% To verify that the measured uncertainties are estimated accurately, we
% computed the \deleted{reduced }$\chi^2$\added{ value} for a linear
% ephemeris fit to the measured TESS mid-transit times.  We found that
% $\chi^2 = 9.2$, with $n=16$ degrees of freedom.  The variance of the
% $\chi^2$ distribution is $2n$, so we would expect $\chi^2 = 16 \pm
% 5.7$.  Visually inspecting the residuals showed that the error
% variance had been overestimated, so we multiplied the measured TESS
% errors by a factor $f=0.76$, forcing a reduced $\chi^2$ of unity.
% This lowered the mean uncertainty of the transit midtimes from $29.8$
% to $22.6$ seconds.  We verified that omitting this step did not
% appreciably alter any of our conclusions.
% 
% Figure~\ref{fig:lightcurves} shows the lightcurve of each individual
% transit, the best-fit models, and the residuals.  Table~2 reports the
% mid-transit times and their uncertainties.  After binning the
% residuals to 1-hour windows, the lightcurves have an rms scatter of
% 586\,{\rm ppm}.  The pre-launch TESS noise
% model\footnote{\url{github.com/lgbouma/tnm}, commit \texttt{be06f09}}
% (\citealt{winn_photonflux_2013}, \citealt{Sullivan_2015} Section 6.4)
% would have predicted an error budget consisting of the following terms
% added in quadrature: 410\,{\rm ppm} from photon-counting noise,
% 202\,{\rm ppm} from detector read noise, and 673\,{\rm ppm} from the
% zodiacal background light.  The level of background light appears to
% have been overestimated in the model.

@@

The periodogram of the final residual (Figure~\ref{fig:splitsignal}
bottom row) shows a weakly significant, poorly resolved peak at
$\approx$8 days, consistent with the visual impression in the time
domain that there could be a weak long-period signal present.
